---
title: "gunsPackage"
author: "Dhuha Manhil"
date: "2022-12-06"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r message = FALSE}
library(gunsPackage)
library(AER)
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(scales)
```

For this final project, I chose to perform statistical analyses on a 
data set called 'Guns'. The data set takes information from 51 states 
(including District of Columbia) and tracks their relevant data over 
23 years, meaning in total there are 1,173 observations, 23 for each 
state. The data was collected from a population of 100,000 in each state. 
Furthermore, the data tracked over a 23 year period, from 1977 to 1999.

There are 13 separate variables that will be used to form some 
conclusions on gun violence and if more guns equals less crime. 
I will be specifically focusing on the state of Virginia, but will 
conduct some additional tests to see how Virginia compares with 
another state on the West Coast.

I chose this specific data set because of my frustration when watching
the news in the United States and seeing more and more shootings 
take place. I would like to see how different gun violence/crime 
was from 1977 to 1999 and compare it to now, seeing if it was less or more.

This package can be used on almost any type of data. 
It doesn't just have to be related to guns or violence, 
rather it can be any data set that contains a wide variety 
of variables. The functions in this package have a general 
purpose of conducting various statistical analyses. 
They should be used to help form comparisons and conclusions 
about the data being studied. Most of the visual functions in 
this package are used to create various visualizations that 
includes bar graphs, histograms, scatter plots, and box plots. 
Some of the other functions in this package deal with statistical 
functions, such as linear regression and correlation.

Reading in the Guns data set.
```{r}
setwd("~/Desktop/gunsPackage/data")
file = "Guns.xlsx"
gunsDataSet <- readingInData(file)
```

The summary function was used here to summarize the data set.
```{r}
summaryGuns <- summarizeFunction(gunsDataSet)
```

Creating and printing a subset to contain only Virginia data.
```{r}
virginiaData <- subset(gunsDataSet, stateid == 46, 
                select = c(year, vio, mur, rob, incarc_rate, pb1064, 
                pw1064, pm1029, pop, avginc, density, stateid, shall))
print(virginiaData)
```

When we run this code below, we can see a summary of the 14
variables that are in the Virginia data set. We can specifically 
focus on the vio, mur, and rob variables, as they can show us 
some descriptive statistics. We can see that the average for vio
is higher than the average for both mur and rob, which can signify
the severity of the level of violence that Virginia endured from 
the years 1977 to 1999.
```{r}
summarizeFunction(virginiaData)
```

Creating a subset of just Virginia Years.
```{r}
virginiaYears <- subset(virginiaData, select = c(year))
print(virginiaYears)
```

Using the bar plot function to produce a graph of the years vs.
the violence crime rate in Virginia.
```{r}
bargraphPlot(virginiaYears, virginiaData$vio, "Violence Crime Rate", 
             "Violence Crime Rate vs Years in the State of Virginia")
```

Creating a new variable that contains the sum of all
crime in Virginia (violence, robbery, and murder).
```{r}
virginiaData$totalCrime <- virginiaData$vio + virginiaData$rob + virginiaData$mur
```

Producing a histogram plot to show frequency of the total
Virginia crime rate.
```{r}
histogramPlot(virginiaData, virginiaData$totalCrime, 
"Total Crime Rate of Murders, Violence, and Robbery", "Frequency", 
"Histogram of Total Crime Rates")
```

Making a scatter plot of the total crime in Virginia vs 
the tncarceration rate.
```{r}
scatterPlot(virginiaData, virginiaData$totalCrime, virginiaData$incarc_rate, "Total Crime",
                "Incarceration Rate", "Total Crime vs Incarceration Rate")
```

Changing the graphics parameters to a window of one row
and two columns of plots.
```{r}
par(mfrow = c(1,2))
```

Plotting a box plot to show the total crime in Virginia
based on if there was a shall law that year.
```{r}
boxPlot(virginiaData, virginiaData$shall == 1, virginiaData$totalCrime, "Shall Law", "Total Crime",
        "Total Crime vs Shall Law")
```
