---
title: "BIOS524FP"
author: "Dhuha Manhil"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(BIOS524FP)
```

This package can be used on almost any type of data. 
It doesn't just have to be related to guns or violence, 
rather it can be any data set that contains a wide variety 
of variables. The functions in this package have a general 
purpose of conducting various statistical analyses. 
They should be used to help form comparisons and conclusions 
about the data being studied. Most of the visual functions in 
this package are used to create various visualizations that 
includes bar graphs, histograms, scatter plots, and box plots. 
Some of the other functions in this package deal with statistical 
functions, such as linear regression and correlation. This package 
and its functions will be used on the R data set ToothGrowth.

The summary function was used here to summarize the data set, ToothGrowth.
```{r}
summaryToothGrowth <- summarizeFunction(ToothGrowth)
```

Here we can see the summary of the data set that includes simple statistical concepts.
```{r}
print(summaryToothGrowth)
```
